---
title: "Visible and NIR Laser Diode Quantities Purchased, 2014-2019"
output:
  flexdashboard::flex_dashboard
    #orientation: columns
    #vertical_layout: fill
runtime: shiny
---

```{r Install Packages, results="hide"}
#if (!require("devtools"))
#  install.packages("devtools")
#devtools::install_github("rstudio/rsconnect")
#install.packages("pacman")
#options(repos="https://cran.rstudio.com" )
library(flexdashboard)
library(tidyverse)
library(grid)
library(readxl)
library(ggplot2)
library(plotly)
#pacman::p_load(rsconnect,flexdashboard,tidyverse,grid,readxl,ggplot2,plotly)
```

Intro
=======================================================================
Column {data-width=1200}
-----------------------------------------------------------------------

This dashboard presents data of a company's sales of visible and NIR laser diodes from 2014 to 2019. Specifically, the aim to determine the most popular quantity of items sold. Items may be purchased individually or at pre-determined "volume pricing" options (i.e., batches of 1, 6, or 11 pieces).

The full distribution of each dataset (visible and NIR) is left skewed, meaning smaller quantites are purchased more frequently than larger ones. Viewing the full distribution means we also see outliers, which stretch all the way to 1500 pieces sold for visible and 2184 pieces sold for NIR. See the two graphs below.

We want to ignore distributions whose tails stretch that far; homing in on the finer details of the distribution from 1-15 is more telling since we specifically want to see if volume pricing options are popular. The "Visible LD Graphs" and "NIR LD Graphs" tabs show more histograms and a bar plot of the quantities purchased in the range of 1-15 pieces.

The bar plots indicate that volume pricing is not especially popular. Diodes sold in quantities of 1 are the most popular, usually followed by the "Other" category, which encompasses quantities that are >1 but <6 as well as >6 but <11; basically, quantities that are between volume pricing batches.

```{r Data Vis}
vis_data <- read_excel("LD_sales_2019.xlsx",sheet="Visible",col_names = TRUE, col_types = NULL, na = "", skip = 0)
vis_data <- vis_data[6:12598,c(1,3)]
colnames(vis_data) <- c("Item", "Quantity")
vis_data$Quantity <- as.numeric(vis_data$Quantity)
# find max and min values of quantities purchased
#max(vis_data$Quantity) # 1500
#min(vis_data$Quantity) # -50
# create separate dataframes for LDs sold in 1, 6, 11 batches and 1, 10, 20, 51, and 101 batches.
#lds <- c("L405G2", "L450G1", "L515A1")
#vis_group1 <- vis_data[!grep1(lds,collapse="|"), vis_data$Item]
# categorize data for bar plot
vis_data$Label <- case_when(
    vis_data$Quantity < 1 ~ as.character("<1"),
    vis_data$Quantity == 1 ~ as.character("1"),
    vis_data$Quantity == 6 ~ as.character("6"),
    vis_data$Quantity == 11 ~ as.character("11"),
    vis_data$Quantity > 11 ~ as.character(">11"),
    TRUE ~ as.character("Other"))
```

```{r Data NIR}
nir_data <- read_excel("LD_sales_2019.xlsx",sheet="NIR",col_names = TRUE, col_types = NULL, na = "", skip = 0)
nir_data <- nir_data[6:7709,c(1,3)]
colnames(nir_data) <- c("Item", "Quantity")
nir_data$Quantity <- as.numeric(nir_data$Quantity)
# find max and min values of quantities purchased
#max(nir_data$Quantity) # 2184
#min(nir_data$Quantity) # -100
# create separate dataframes for LDs sold in 1, 6, 11 batches and 1, 10, 20, 51, and 101 batches.
#lds2 <- c("L785H1")
#nir_group1 <- nir_data[!grep1(lds2,collapse="|"), nir_data$Item]
# categorize data for bar plot
nir_data$Label <- case_when(
    nir_data$Quantity < 1 ~ as.character("<1"),
    nir_data$Quantity == 1 ~ as.character("1"),
    nir_data$Quantity == 6 ~ as.character("6"),
    nir_data$Quantity == 11 ~ as.character("11"),
    nir_data$Quantity > 11 ~ as.character(">11"),
    TRUE ~ as.character("Other"))

#hist(vis_data$Quantity, breaks=100, xlab = "Qty. Purchased",main = paste("Full Distribution of Visible LDs"))
#hist(nir_data$Quantity, breaks=100, xlab = "Qty. Purchased",main = paste("Full Distribution of NIR LDs"))
```

```{r }
p1 = ggplot(vis_data, aes(vis_data$Quantity)) +geom_histogram() + labs(title="Full Distribution of Visible LDs", x = "Qty. Purchased", y = "Frequency")
p2 = ggplot(nir_data, aes(nir_data$Quantity)) +geom_histogram() + labs(title="Full Distribution of NIR LDs",x = "Qty. Purchased", y = "Frequency")

ggplotly(p1, width = 700, height = 400)
ggplotly(p2, width = 700, height = 400)
```

```{r }

```

Visible LD Graphs
=======================================================================
Column {.sidebar}
-----------------------------------------------------------------------

<b>Histogram tab:</b><br />
Use the tools below to change the number of bins as well as the bandwidth, which is simply a curve that fits the histogram. The histogram shows distribution of all visible LDs where the quantity purchased is between 1 and 15. For full distributions, take a look at the Individual Histograms tab.
```{r}
selectInput("n_breaks", label = "Number of bins:",
            choices = c(1, 3, 5, 10, 15), selected = 5)

sliderInput("bw_adjust", label = "Bandwidth adjustment:",
            min = 1, max = 2, value = 1.5, step = 0.1)
```

<b>Individual Histograms tab:</b><br />
These individual histograms (organized in a lattice form to display all item #s) show the distribution of LD quantities sold from 1 to 15. Note that item #s L405G2, L450G1, L515A1, and L785H1 offer volume pricing for 1, 10, 20, 51, and 101 pieces.

<b>Bar Plots tab:</b><br />
The bar plots (again, in a lattice to display all item #s) count the number of LDs purchased in one of 6 groups: <1, 1, 6, 11, >11, and "Other". The "Other" category simply catches quantities purchased between the volume pricing options, i.e., it contains the number of LDs purchased (x) in these quantities:
1 < x < 6
7 < x < 11

row {.tabset }
-----------------------------------------------------------------------

### Histogram

```{r}
renderPlot({
  with(vis_data, hist(Quantity[Quantity >= 1 & Quantity <= 15], probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Qty. Purchased", main = "Histogram of All Visible LDs, Qty. Purchased b/w 1 and 15"))
  
  dens <- density(vis_data$Quantity, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

### Individual Histograms
```{r Histograms Vis}
# visible LD histograms
#histogram(~Quantity|Item,data=vis_data, type="count", xlab="Qty. Purchased", main="Individual Histograms of Qty. Purchased", breaks=c(6))
vis_hist <- qplot(Quantity, facets = ~ Item, data = vis_data, geom="histogram", ylim=c(0,200), binwidth = 0.5, main="Individual Histograms of Qty. Purchased", xlab = "Qty. Purchased", ylab = "Frequency") + theme(plot.title = element_text(hjust = 0.5)) + scale_x_continuous(breaks = seq(0, 15, 5), lim = c(0,15))# +theme(axis.text=element_text(size=4))
ggplotly(vis_hist)
```



### Bar Plots
```{r Bar Plots}
positions <- c("<1", "1", "6", "11", ">11","Other") # write positions of labels on x-axis

# visible LDs bar plot
vis_bar <- qplot(Label, facets = ~ Item, data = vis_data,geom="bar", main="Individual Bar Plots of Qty. Purchased", xlab = "Qty. Purchased", ylab = "Count") + theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(limits = positions)+ scale_y_continuous(breaks = seq(0, 400, 100), lim = c(0,400))# +theme(axis.text=element_text(size=4))
ggplotly(vis_bar)

```

NIR LD Graphs
=======================================================================
Column {.sidebar}
-----------------------------------------------------------------------
### NIRLaser Diodes


<b>Histogram tab:</b><br />
Use the tools below to change the number of bins as well as the bandwidth, which is simply a curve that fits the histogram. The histogram shows distribution of all visible LDs where the quantity purchased is between 1 and 15. For full distributions, take a look at the Individual Histograms tab.

```{r}
selectInput("n_breaks", label = "Number of bins:",
            choices = c(1, 3, 5, 10, 15), selected = 5)

sliderInput("bw_adjust", label = "Bandwidth adjustment:",
            min = 1, max = 2, value = 1.5, step = 0.1)
```
<b>Individual Histograms tab:</b><br />
The individual histograms show the full distribution of LD quantities sold, limited from 1 to 15. If outliers are included, then there will only be one large bar that spans 0-50, which doesn't provide enough detail. Note that item #s L405G2, L450G1, L515A1, and L785H1 offer volume pricing for 1, 10, 20, 51, and 101 pieces.

<b>Bar Plots tab:</b><br />
These bar plots show the frequency of quantities purchased, categorized into 6 groups: <1, 1, 6, 11, >11, and Other. The Other category contains the number of LDs purchased (x) in these quantities:
1 < x < 6
7 < x < 11

row {.tabset }
-----------------------------------------------------------------------

### Histogram

```{r}
renderPlot({
  with(nir_data, hist(Quantity[Quantity >= 1 & Quantity <= 15], probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Qty. Purchased", main = "Histogram of All NIR LDs, Qty. Purchased b/w 1 and 15"))
  
  dens <- density(nir_data$Quantity, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

### Individual Histograms

```{r Histograms NIR}
# NIR LD histograms
#histogram(~Quantity|Item,data=nir_data, type="count", xlab="Qty. Purchased", main="Individual Histograms of Qty. Purchased", breaks=c(6))

nir_hist <- qplot(Quantity, facets = ~ Item, data = nir_data, geom="histogram", xlim=c(1,15), ylim=c(0,150), binwidth = 0.5, main="Individual Histograms of Qty. Purchased", xlab = "Qty. Purchased", ylab = "Frequency") + theme(plot.title = element_text(hjust = 0.5))+ scale_x_continuous(breaks = seq(0, 15, 5), lim = c(0,15))# +theme(axis.text=element_text(size=4))
ggplotly(nir_hist)
```

### Bar Plots

```{r Bar Plots NIR}
positions <- c("<1", "1", "6", "11", ">11","Other") # write positions of labels on x-axis

# NIR LDs bar plot
nir_bar <- qplot(Label, facets = ~ Item, data = nir_data,geom="bar", main="Individual Bar Plots of Qty. Purchased", xlab = "Qty. Purchased", ylab = "Count") + theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(limits = positions)+ scale_y_continuous(breaks = seq(0, 400, 100), lim = c(0,400))#+theme(axis.text=element_text(size=4))
ggplotly(nir_bar)
```
